$\pi$SOD-M: A Methodology for Building Service-Oriented Applications in the Presence of Non-Functional Properties

### 2013-12-16 JournalToSubmit2013/relworks.tex
main y related works par genoveva

Related works to our approach includes existing proposal dealing with the inclusion of non-functional properties in service-oriented development. Standards devoted for expressing non-functional properties for service-oriented applications are also analyzed in this section which needs to be addresses in the code generation phase. Finally exiting methodology for the development of service-oriented applications including non-functional properties are here depicted and compared with our approach.
%As a result of our study of related works, in this section we present a categorization of main concepts related the treatment if non-functional requirement in the field of service-oriented development.

@@

In previous work we have carried out a systematic review \cite{placidoPhDThesis2012} in order to analyze how non-functional properties are taking into account in existing approaches for service-oriented development. Existing proposals in this field may be classified in three types of works which are described in the followed.

The first types of works are thus considering a particular non-functional concern (e.g. security) which is modeled and then "glued" or associated to functional models of an application. The work of Chollet et al. \cite{CholletL09} defines a proposal to associate non-functional quality properties (security properties in that case) to functional activities which are modeled in a web service composition model.
Schmeling et al. \cite{SchmelingCM11} present an approach and also a toolset for specifying and implementing non-functional concerns in web service compositions; non-functional concerns are modeled first in non-functional model and then related to a service composition represented in a BPMN diagram. Ovaska et al. \cite{OvaskaEHPA10} present an approach to support quality management at design time. Quality requirements are modeled in a first phase and represented then in an architectural model where quality requirements are associated to some components in the model. In our case we propose to model non-functional properties joint to the modeling of functional concerns what simplifies the web service development process. Non-functional properties are represented in our work as constraints and policies but more general quality model such as the proposed by \cite{Goeb2011,Klass2009} could be taken in consideration in further works.

Second type of works related to our proposal proposes specifics architecture or frameworks to manage and validate QoS attributes in service composition processes. The work of Xiao et al. \cite{XiaoCZBOLH08} present a framework to verify SLA compliance in composed services. Task in a BPEL process can be annotated with non-functional attributes (such as time, cost, resources, etc.) which are validated at design time. A similar work is presented in \cite{SchmelingCM11} by using a formal method that allows to choice best suited services instances according to a service composition pattern that include some QoS attributes. Babamir et al. present in \cite{Babamir2010} a framework for evaluate quality attributes of web services. By executing an algorithm, the framework allow to choice best web services for any task in a process and also candidate web service to replace original ones in cases it cannot perform at runtime. In \cite{Karunamurthy2012787} authors propose architecture in which web services may be described with non-functional characteristics that are validated in the service composition process.

@@

>>>>

We conducted a systematic review \cite{placidoPhDThesis2012} in order to analyze approaches that consider non-functional properties for developing service-oriented applications. These  approaches address the problem from three perspectives: NFP expressed as quality of service properties (QoS attributes) that can be used for looking up services that can participate in a service composition, programming NFP as protocols that are asserted to service composition code, and methodologies that provide tools for building service compositions with associated NFP.

%  Standards devoted for expressing non-functional properties for service-oriented applications are also analyzed in this section which needs to be addresses in the code generation phase. Finally exiting methodology for the development of service-oriented applications including non-functional properties are here depicted and compared with our approach.
%As a result of our study of related works, in this section we present a categorization of main concepts related the treatment if non-functional requirement in the field of service-oriented development.

@@

The first types of works are thus considering a particular non-functional concern (e.g. security) which is modeled and then "glued" or associated to functional models of an application.

\cite{CholletL09} defines a proposal to associate non-functional quality properties (security properties in that case) to functional activities which are modeled in a web service composition model.

\cite{SchmelingCM11} present an approach and also a toolset for specifying and implementing non-functional concerns in web service compositions; non-functional concerns are modeled first in non-functional model and then related to a service composition represented in a BPMN diagram.

 \cite{OvaskaEHPA10} present an approach to support quality management at design time. Quality requirements are modeled in a first phase and represented then in an architectural model where quality requirements are associated to some components in the model.

Second type of works related to our proposal proposes specifics architecture or frameworks to manage and validate QoS attributes in service composition processes.

\cite{XiaoCZBOLH08} present a framework to verify SLA compliance in composed services. Task in a BPEL process can be annotated with non-functional attributes (such as time, cost, resources, etc.) which are validated at design time. A similar work is presented in \cite{SchmelingCM11} by using a formal method that allows to choice best suited services instances according to a service composition pattern that include some QoS attributes. Babamir et al. present in \cite{Babamir2010} a framework for evaluate quality attributes of web services. By executing an algorithm, the framework allow to choice best web services for any task in a process and also candidate web service to replace original ones in cases it cannot perform at runtime. In \cite{Karunamurthy2012787} authors propose architecture in which web services may be described with non-functional characteristics that are validated in the service composition process.

@@

In our case we propose to model non-functional properties joint to the modeling of functional concerns what simplifies the web service development process. Non-functional properties are represented in our work as constraints and policies but more general quality model such as the proposed by \cite{Goeb2011,Klass2009} could be taken in consideration in further works.
